\chapter{General Relativity}
\label{chap:general-relativity}

This chapter summarizes the parts of general relativity that are needed in the
rest of the thesis. We work in geometrized units \(G=c=1\) and use the metric
signature \((-+++)\). Greek indices
indicate space-time coordinate components and we follow Einstein’s summation
rule over the same-index pairs (unless they denote a specific coordinate). The systems studied below are free test particle systems.

\section{Geodesic Motion}
\label{sec:geodesic-motion}

A space-time is a four-dimensional Lorentzian manifold with metric tensor
\(g_{\mu\nu}\). In local coordinates \(x^\mu\), the line element is
\begin{equation}
    \mathrm{d}s^2
    =
    g_{\mu\nu}\,\mathrm{d}x^\mu\mathrm{d}x^\nu .
    \label{eq:line-element}
\end{equation}
For a massive free test particle, the world-line can be parametrized by its
proper time \(\tau\). Its four-velocity is
\begin{equation}
    u^\mu
    =
    \frac{\mathrm{d}x^\mu}{\mathrm{d}\tau},
    \label{eq:four-velocity}
\end{equation}
and the normalization of a timelike world-line is
\begin{equation}
    g_{\mu\nu}u^\mu u^\nu=-1.
    \label{eq:timelike-normalization-chap2}
\end{equation}
The free motion is a geodesic motion. In coordinates, the geodesic equation is
\begin{equation}
    \frac{\mathrm{d}^2x^\mu}{\mathrm{d}\tau^2}
    +
    \Gamma^\mu_{\alpha\beta}
    \frac{\mathrm{d}x^\alpha}{\mathrm{d}\tau}
    \frac{\mathrm{d}x^\beta}{\mathrm{d}\tau}
    =
    0,
    \label{eq:geodesic-equation}
\end{equation}
where the Christoffel symbols are determined by the metric,
\begin{equation}
    \Gamma^\mu_{\alpha\beta}
    =
    \frac{1}{2}g^{\mu\nu}
    \left(
        g_{\nu\alpha,\beta}
        +
        g_{\beta\nu,\alpha}
        -
        g_{\alpha\beta,\nu}
    \right).
    \label{eq:christoffel-symbols}
\end{equation}
\cite{BicakSemerak2026}

\section{Symmetries and Conserved Quantities}
\label{sec:symmetries-conserved-quantities}

The space-times used in this thesis are stationary and axially symmetric. Therefore, two Killing vector fields exist, one associated with time
translations and one with rotations around the symmetry axis. In adapted
coordinates, we denote them by
\begin{equation}
    \xi^\mu_{(t)}=\delta^\mu_t,
    \qquad
    \xi^\mu_{(\phi)}=\delta^\mu_\phi .
    \label{eq:killings}
\end{equation}
If \(\xi^\mu\) is a Killing vector field, then
\(g_{\mu\nu}u^\mu\xi^\nu\) is conserved along a geodesic. For the two
symmetries above, this gives the specific energy and azimuthal angular momentum
\begin{equation}
    \epsilon
    :=
    -g_{\mu\nu}u^\mu\xi^\nu_{(t)},
    \qquad
    \ell
    :=
    g_{\mu\nu}u^\mu\xi^\nu_{(\phi)}.
    \label{eq:conservation}
\end{equation}
For a diagonal stationary-axisymmetric metric, this means
\begin{equation}
    u^t=-\frac{\epsilon}{g_{tt}},
    \qquad
    u^\phi=\frac{\ell}{g_{\phi\phi}}.
    \label{eq:ut-uphi-conserved}
\end{equation}

\section{Weyl Space-times}
\label{sec:weyl-spacetimes}

Static axisymmetric space-times that satisfy the energy-momentum tensor condition $T_\rho^\rho + T^z_z$ can be written in Weyl cylindrical-type coordinates
\((t,\phi,\rho,z)\) \cite{Basovnik2010}. The Weyl class solutions can then be described by the metric
\begin{equation}
    \mathrm{d}s^2
    =
    -e^{2\nu}\,\mathrm{d}t^2
    +
    \rho^2 e^{-2\nu}\,\mathrm{d}\phi^2
    +
    e^{2(\lambda-\nu)}
    \left(
        \mathrm{d}\rho^2+\mathrm{d}z^2
    \right),
    \label{eq:weyl-metric}
\end{equation}
where \(\nu=\nu(\rho,z)\) and \(\lambda=\lambda(\rho,z)\). Equivalently,
with \(N\equiv e^\nu\),
\begin{equation}
    \mathrm{d}s^2
    =
    -N^2\,\mathrm{d}t^2
    +
    N^{-2}
    \left[
        \rho^2\,\mathrm{d}\phi^2
        +
        e^{2\lambda}
        \left(
            \mathrm{d}\rho^2+\mathrm{d}z^2
        \right)
    \right].
    \label{eq:weyl-metric-lapse}
\end{equation}
The first metric function satisfies the axisymmetric Laplace
equation,
\begin{equation}
    \nu_{,\rho\rho}
    +
    \frac{1}{\rho}\nu_{,\rho}
    +
    \nu_{,zz}
    =
    0.
    \label{eq:weyl-laplace}
\end{equation}
In the weak-field
limit, \(g_{tt}=-e^{2\nu}\approx -(1+2\nu)\), so \(\nu\) has the role of the
Newtonian gravitational potential. The second metric function is fixed by
\begin{align}
    \lambda_{,\rho}
    &=
    \rho
    \left[
        \left(\nu_{,\rho}\right)^2
        -
        \left(\nu_{,z}\right)^2
    \right],
    \label{eq:weyl-lambda-rho}
    \\
    \lambda_{,z}
    &=
    2\rho\nu_{,\rho}\nu_{,z}.
    \label{eq:weyl-lambda-z}
\end{align}
The equations for \(\lambda\) are nonlinear in \(\nu\). Therefore, $\nu$ can be added linearly, but the corresponding \(\lambda\) of a
superposed configuration contains interaction terms.
\cite{Sychrovsky2020,Kraus2025,Basovnik2010}

\section{Schwarzschild Black Hole}
\label{sec:schwarzschild-weyl}

The Schwarzschild metric in standard coordinates \((t,r,\theta,\phi)\) is
\begin{equation}
    \mathrm{d}s^2
    =
    -\left(1-\frac{2M}{r}\right)\mathrm{d}t^2
    +
    \left(1-\frac{2M}{r}\right)^{-1}\mathrm{d}r^2
    +
    r^2
    \left(
        \mathrm{d}\theta^2
        +
        \sin^2\theta\,\mathrm{d}\phi^2
    \right).
    \label{eq:schwarzschild-metric}
\end{equation}
It describes the exterior of a nonrotating, charge-less black hole of mass \(M\). The
transformation to Weyl coordinates is
\begin{equation}
    \rho
    =
    \sqrt{r(r-2M)}\sin\theta,
    \qquad
    z
    =
    (r-M)\cos\theta.
    \label{eq:schwarzschild-to-weyl}
\end{equation}
\cite{Semerak2016}
If we denote

\begin{equation}
    d_{\pm}
    :=
    \sqrt{\rho^2+(z\pm M)^2},
    \label{eq:schwarzschild-distances}
\end{equation}

the metric functions $\nu$ and $\lambda$ of the Schwarzschild solution are
\begin{align}
    \nu_{\mathrm{Schw}}
    &=
    \frac{1}{2}
    \ln
    \left(
        \frac{d_{+}+d_{-}-2M}
             {d_{+}+d_{-}+2M}
    \right),
    \label{eq:schwarzschild-potential}
    \\
    \lambda_{\mathrm{Schw}}
    &=
    \frac{1}{2}
    \ln
    \left[
        \frac{(d_{+}+d_{-})^2-4M^2}
             {4d_{+}d_{-}}
    \right].
    \label{eq:schwarzschild-lambda}
\end{align}
\cite{Kraus2025}.

\section{Bach-Weyl Ring}
\label{sec:bach-weyl-ring}

The Bach-Weyl solution represents a static, axially symmetric,
infinitesimally thin circular ring. We denote by \(b\) the Schwarzschild
radial coordinate of the ring. Its Weyl
cylindrical radius is
\begin{equation}
    \sigma_{\mathrm{BW}}
    =
    \sqrt{b(b-2M)},
    \label{eq:schw-bw-ring-radius}
\end{equation}
which follows from the Schwarzschild--Weyl transformation
\eqref{eq:schwarzschild-to-weyl} at \(\theta=\pi/2\). In the ring potential
below, \(\sigma\) denotes this Weyl radius. Let \(\mathcal{M}\) be the ring
mass and define
\begin{equation}
    l_{1,2}
    :=
    \sqrt{(\rho\mp \sigma)^2+z^2},
    \qquad
    k^2
    :=
    1-\frac{l_1^2}{l_2^2}
    =
    \frac{4\sigma\rho}{l_2^2}.
    \label{eq:ring-quantities}
\end{equation}
The complete elliptic integral of the first kind is
\begin{equation}
    K(k)
    :=
    \int_0^{\pi/2}
    \frac{\mathrm{d}\alpha}
         {\sqrt{1-k^2\sin^2\alpha}}.
    \label{eq:complete-elliptic-k}
\end{equation}
The Bach-Weyl potential is
\begin{equation}
    \nu_{\mathrm{BW}}
    =
    -\frac{2\mathcal{M}K(k)}{\pi l_2}.
    \label{eq:bw-potential}
\end{equation}

It is the relativistic counterpart of the Newtonian potential of a homogeneous
thin ring. However, the function \(\lambda_{\mathrm{BW}}\) does not have a Newtonian counterpart.
It is obtained from Eqs.~\eqref{eq:weyl-lambda-rho} and
\eqref{eq:weyl-lambda-z}. Close to the ring,
the behavior of $\lambda$ makes the space properties very different from those of the Newtonian case. In particular,
the space is strongly anisotropic rather than locally cylindrical around the ring.
This makes the Bach-Weyl ring geometrically more severe than its Newtonian
analog \cite{Semerak2016}.

Since Eq.~\eqref{eq:weyl-laplace} is linear, the Schwarzschild and
Bach-Weyl potentials can be superposed
\begin{equation}
    \nu
    =
    \nu_{\mathrm{Schw}}
    +
    \nu_{\mathrm{BW}}.
    \label{eq:schwarzschild-bw-potential}
\end{equation}

The second Weyl function is not a simple sum. It has the structure
\begin{equation}
    \lambda
    =
    \lambda_{\mathrm{Schw}}
    +
    \lambda_{\mathrm{BW}}
    +
    \lambda_{\mathrm{int}},
    \label{eq:schwarzschild-bw-lambda}
\end{equation}
where \(\lambda_{\mathrm{int}}\) is the interaction contribution generated by
non-linear equations for \(\lambda\) \cite{Basovnik2010}. The combined metric
Eqs.~\eqref{eq:weyl-metric} and
\eqref{eq:schwarzschild-bw-potential} is the Schwarzschild-Bach-Weyl
(SCHW+BW) space-time used in the first part of the numerical study. 

In
the following chapters, the abbreviation SCHW+BW always refers to this
Schwarzschild black hole plus Bach-Weyl ring system; SCHW denotes the
Schwarzschild black hole and BW denotes the Bach-Weyl ring.

\section{Majumdar-Papapetrou Space-times}
\label{sec:majumdar-papapetrou}

Majumdar-Papapetrou space-times are static electrovacuum solutions. The Majumdar-Papapetrou solutions are a subclass of the Weyl class solutions. Hence, we refer to them as a Majumdar-Papapetrou subclass. In cylindrical coordinates they can be written as
\begin{equation}
    \mathrm{d}s^2
    =
    -N^2\,\mathrm{d}t^2
    +
    N^{-2}
    \left(
        \rho^2\,\mathrm{d}\phi^2
        +
        \mathrm{d}\rho^2
        +
        \mathrm{d}z^2
    \right).
    \label{eq:mp-metric}
\end{equation}
This is a special case of the Weyl form with \(\lambda=0\). For several sources, one has a linear superposition in
\(N^{-1}\), as it satisfies the axisymmetric Laplace equation \cite{Kraus2025,Polcar2019}.

\subsection{Extreme Reissner-Nordstr\"om Black Hole}
\label{subsec:extreme-rn}

The Reissner-Nordstr\"om metric in spherical coordinates, with the same meaning as in the Schwarzschild case, is
\begin{equation}
    \mathrm{d}s^2
    =
    -f(r)\,\mathrm{d}t^2
    +
    f(r)^{-1}\,\mathrm{d}r^2
    +
    r^2
    \left(
        \mathrm{d}\theta^2
        +
        \sin^2\theta\,\mathrm{d}\phi^2
    \right),
    \label{eq:rn-metric}
\end{equation}
where
\begin{equation}
    f(r)
    =
    1-\frac{2M}{r}+\frac{Q^2}{r^2}.
    \label{eq:rn-function}
\end{equation}
For the extreme case, \(|Q|=M\), we obtain
\begin{equation}
    f(r)
    =
    \left(1-\frac{M}{r}\right)^2.
\end{equation}
The corresponding Weyl-type coordinates satisfy
\begin{equation}
    \rho=(r-M)\sin\theta,
    \qquad
    z=(r-M)\cos\theta.
    \label{eq:ern-weyl-coordinates}
\end{equation}
The extreme
Reissner-Nordstr\"om lapse can be written as
\begin{equation}
    \frac{1}{N_{\mathrm{eRN}}}
    =
    1+\frac{M}{\sqrt{\rho^2+z^2}}.
    \label{eq:ern-lapse}
\end{equation}
\cite{Semerak2016, Kraus2025, BicakSemerak2026}

\subsection{Majumdar-Papapetrou Ring}
\label{subsec:mp-ring}

A solution consisting of a number
of identical extreme black holes arranged in a circle with a continuous limit of this situation when one has an
infinite number of “infinitesimal” holes distributed along
a circular ring defines the Majumdar-Papapetrou ring. We again use \(b\) for the radial coordinate of the ring used in
the corresponding Reissner-Nordstr\"om coordinates. Comparing $g_{tt}$ for the extreme
Reissner-Nordstr\"om in spherical coordinates \(g_{tt}=-(1-\frac{M}{r})^2\) with the form in isotropic coordinates \(g_{tt}=-N^2\), \(N^{-1}=1+\frac{M}R{}\), gives
\(r=R+M\). As Cartesian-type coordinates relate to Weyl coordinates by $x = \rho\sin\phi$, $y = \rho\cos\phi$ and $R^2\equiv x^2+y^2+z^2$, we get $R^2 = \rho^2 + z^2=(r-M)^2$. Since the disk is in the $z=0$ plane, we obtain $R=\rho=r-M$.
Thus, a ring at \(r=b\) is represented in the
Majumdar-Papapetrou cylindrical coordinates by
\begin{equation}
    \sigma_{\mathrm{MP}}
    =
    b-M.
    \label{eq:rn-mp-ring-radius}
\end{equation}
For a ring of cylindrical radius \(\sigma\) and mass \(\mathcal{M}\), the
inverse lapse is
\begin{equation}
    \frac{1}{N_{\mathrm{MP}}}
    =
    1
    +
    \frac{2\mathcal{M}K(k)}{\pi l_2},
    \label{eq:mp-ring-lapse}
\end{equation}
with \(l_1,l_2\) and \(k\) defined by Eq.~\eqref{eq:ring-quantities}, using
\(\sigma=\sigma_{\mathrm{MP}}\). In
contrast to the Bach-Weyl ring, the Majumdar-Papapetrou ring has
\(\lambda=0\).
\cite{Semerak2016,Polcar2019}.

Because both the extreme Reissner-Nordstr\"om black hole and the
Majumdar-Papapetrou ring belong to the same class of solutions, the combined
Reissner-Nordstr\"om--Majumdar-Papapetrou (RN+MP) system is obtained by
adding their contributions to \(N^{-1}\)
\begin{equation}
    \frac{1}{N}
    =
    1
    +
    \frac{M}{\sqrt{\rho^2+z^2}}
    +
    \frac{2\mathcal{M}K(k)}{\pi l_2}.
    \label{eq:ern-mp-superposition}
\end{equation}
Together with Eq.~\eqref{eq:mp-metric}, this gives the RN+MP metric used in
this thesis. In the following chapters, RN+MP denotes the
extreme Reissner-Nordstr\"om black hole plus Majumdar-Papapetrou ring
system, RN denotes the extreme Reissner-Nordstr\"om and MP denotes
Majumdar-Papapetrou.