%%% Please fill in basic information on your thesis, which will be automatically
%%% inserted at the right places. You need to replace \xxx{...} by real data.

% Type of your thesis:
%	"bc" for Bachelor's
%	"mgr" for Master's
%	"phd" for PhD
%	"rig" for rigorosum
\def\ThesisType{bc}

% Language of your study programme:
%	"cs" for Czech
%	"en" for English
\def\StudyLanguage{cs}

% Thesis title in English (exactly as in the official assignment)
% (Note: \xxx is a "ToDo label" which makes the unfilled visible. Remove it.)
\def\ThesisTitle{Fractal features of geodesic motion around perturbed black holes}

% Author of the thesis (you)
\def\ThesisAuthor{Dávid Jopek}

% Year when the thesis is submitted
\def\YearSubmitted{2026}

% Name of the department or institute, where the work was officially assigned
% (according to the Organizational Structure of MFF UK in English,
% see https://www.mff.cuni.cz/en/faculty/organizational-structure,
% or a full name of a department outside MFF)
\def\Department{Institute of Theoretical Physics}

% Is it a department (katedra), or an institute (ústav)?
\def\DeptType{Institute}

% Thesis supervisor: name, surname and titles
\def\Supervisor{doc. RNDr. Old\v{r}ich Semer\'ak, DSc.}

% Supervisor's department (again according to Organizational structure of MFF)
\def\SupervisorsDepartment{Institute of Theoretical Physics}

% Study programme (does not apply to rigorosum theses)
\def\StudyProgramme{Physics}

% An optional dedication: you can thank whomever you wish (your supervisor,
% consultant, who provided you with tea and pizza, etc.)
\def\Dedication{%
I dedicate this thesis to all the people who make my life great, help me love what I do, and find joy in it, as well as to those who support me on my journey and help me become a better person.
}

% Abstract (recommended length around 80-200 words; this is not a copy of your thesis assignment!)
\def\Abstract{%
We study the dynamics of free test particles in the gravitational field of a Schwarzschild and a Reissner-Nordstr\"om black hole perturbed by an external exact thin axisymmetric source. Specifically, we consider a Schwarzschild black hole with a Bach-Weyl ring and a Reissner-Nordstr\"om black hole with a Majumdar-Papapetrou ring. To quantify chaotic dynamics, we construct basin maps and estimate the fractal dimension of basin boundaries using the uncertainty exponent method. We reproduce and compare selected results available in the literature and extend the basin map analysis to parameter regimes and configurations not previously studied with this approach.
}

% 3 to 5 keywords (recommended) separated by \sep
% Keywords are useful for indexing and searching for the theses by topic.
\def\ThesisKeywords{%
general relativity\sep perturbed black holes\sep geodesic chaos\sep basin boundaries\sep uncertainty exponent
}

% If any of your metadata strings contains TeX macros, you need to provide
% a plain-text version for use in XMP metadata embedded in the output PDF file.
% If you are not sure, check the generated thesis.xmpdata file.
% Plain-text versions for XMP metadata (avoid TeX macros / diacritics here).
\def\ThesisAuthorXMP{David Jopek}
\def\ThesisTitleXMP{Fractal features of geodesic motion around perturbed black holes}
\def\ThesisKeywordsXMP{general relativity; perturbed black holes; geodesic chaos; basin boundaries; uncertainty exponent}
\def\AbstractXMP{We study the dynamics of free test particles in the gravitational field of a Schwarzschild and a Reissner-Nordstrom black hole perturbed by an external exact thin axisymmetric source. Specifically, we consider a Schwarzschild black hole with a Bach-Weyl ring and a Reissner-Nordstrom black hole with a Majumdar-Papapetrou ring. To quantify chaotic dynamics, we construct basin maps and estimate the fractal dimension of basin boundaries using the uncertainty exponent method. We reproduce and compare selected results available in the literature and extend the basin map analysis to parameter regimes and configurations not previously studied with this approach.}

% If your abstracts are long and do not fit in the infopage, you can make the
% fonts a bit smaller by this setting. (Also, you should try to compress your abstract more.)
\def\InfoPageFont{}
%\def\InfoPageFont{\small}  % uncomment to decrease font size

% If you are studing in a Czech programme, you also need to provide metadata in Czech:
% (in English programmes, this is not used anywhere)

\def\ThesisTitleCS{Fraktální aspekty geodetického pohybu v okolí perturbovaných černých děr}
\def\DepartmentCS{Ústav teoretické fyziky}
\def\DeptTypeCS{Ústav}
\def\SupervisorsDepartmentCS{Ústav teoretické fyziky}
\def\StudyProgrammeCS{Fyzika}

\def\ThesisKeywordsCS{%
obecná relativity\sep porušené černé díry\sep geodetický chaos\sep hranice oblastí přitažlivosti\sep fraktální dimenze\sep exponent neurčitosti
}

\def\AbstractCS{%
Studujeme dynamiku volných testovacích částic v gravitačním poli
Schwarzschildovy a Reissnerovy-Nordströmovy černé díry perturbované vnější
exaktní tenkou osově symetrickou soustavou. Konkrétně uvažujeme
Schwarzschildovu černou díru s Bachovým-Weylovým prstencem a
Reissnerovu-Nordströmovu černou díru s Majumdarovým-Papapetrouovým
prstencem. Pro kvantifikaci chaotické dynamiky konstruujeme mapy oblastí
přitažlivosti a odhadujeme fraktální dimenzi hranic těchto oblastí metodou
exponentu neurčitosti. Reprodukujeme a porovnáváme vybrané výsledky dostupné v
literatuře a rozšiřujeme analýzu map oblastí přitažlivosti na režimy parametrů
a konfigurace, které dosud nebyly tímto přístupem studovány.
}
